A standardized framework for risk-based assessment of treatment effect heterogeneity in observational healthcare databases

### 2021-09-02 submission/manuscript.rmd
remove evaluation section and distribute to prediction and estimation sections

***Cohort definitions are crucial for this step of the framework. We define a
cohort as the set of patients who satisfy one or more inclusion criteria for a
duration of time. A cohort within the OHDSI setting is more than a set of
specific clinical codes, providing a definition of a logic for how to use that
code set. All cohort definitions consist of: an entry event, i.e. the time a
patient enters a cohort; a set of inclusion criteria applied to the initial
event cohort to further restrict the set of people, resulting in the
construction of the construction of the qualifying cohort; cohort exit criteria
that terminate the patient's presence in the cohort. Cohort definitions are
transportable, meaning they can be implemented in any database, provided that it
is mapped to the OMOP-CDM.***

@@

@@

## Step 5: Evaluation

In each risk stratum we need to evaluate if adequate covariate balance was
achieved accounting for measured confounding. If that is not the case,
interpretation of the results may be problematic. Common approaches include
evaluation of the overlap of propensity score distributions and calculation of
standardized covariate differences before and after propensity score adjustment.

A performance overview of the derived prediction models, including
discrimination and calibration both in the propensity score matched subset, the
entire population and separately for treated and comparator patients should also
be reported. This is quite relevant as the performance of the prediction models
is directly related to our ability to single out patient subgroups where
treatment may be highly beneficial or unsafe. Kent et al [@Kent2016]
demonstrated that the event rate and the discriminative ability of the
prediction model can predict very well the distribution of predicted risk. Lower
event rate and higher c-statistic (given good calibration) result in high risk
heterogeneity, thus making estimated average treatment effects uninformative. In
this case, risk stratified analysis of HTE can be more effective in singling out
patient subgroups that stand to benefit (or be harmed) most by treatment in
question.

## Step 6: Presentation of results

>>>>

***Cohort definitions are crucial for this step of the framework. We define a
cohort as the set of patients who satisfy one or more inclusion criteria for a
duration of time. A cohort within the OHDSI setting is more than a set of
specific clinical codes, providing a definition of a logic for how to use that
code set. All cohort definitions consist of: an entry event, i.e. the time a
patient enters a cohort; a set of inclusion criteria applied to the initial
event cohort to further restrict the set of people, resulting in the
construction of the construction of the qualifying cohort; cohort exit criteria
that terminate the patient's presence in the cohort. Cohort definitions are
transportable, meaning that in theory they can be implemented in any database,
provided that it is mapped to the OMOP-CDM.***

@@

After model development, a performance overview of the derived prediction models
including discrimination and calibration both in the propensity score matched
subset, the entire population and separately for treated and comparator patients
should also be reported. This is important to ensure that no overfitting of the
prediction model in one of the cohorts has occurred. In addition, the
performance of the prediction models is directly related to our ability to
single out patient subgroups where treatment may be highly beneficial or unsafe.
Kent et al [@Kent2016] demonstrated that the event rate and the discriminative
ability of the prediction model can predict very well the distribution of
predicted risk. Lower event rate and higher c-statistic (given good calibration)
result in high risk heterogeneity, thus making estimated average treatment
effects uninformative. In this case, risk stratified analysis of HTE can be more
effective in singling out patient subgroups that stand to benefit (or be harmed)
most by treatment in question.

@@

Before focusing on the results of the estimation process we need to evaluate if
adequate covariate balance was achieved within each risk stratum accounting for
measured confounding. Common approaches include evaluation of the overlap of
propensity score distributions and calculation of standardized covariate
differences before and after propensity score adjustment.

## Step 5: Presentation of results
